The SMT Competition 2015--2018

### 2019-08-31 2018/report/SMT-COMP.tex
Layout (avoid overfull hboxes)
Layout: increase floatpagefraction, \include -> \input, \% -> \percent

% float layout
\setcounter{topnumber}{2}
\setcounter{bottomnumber}{2}
\setcounter{totalnumber}{4}
\renewcommand{\topfraction}{0.85}
\renewcommand{\bottomfraction}{0.85}
\renewcommand{\textfraction}{0.15}
\renewcommand{\floatpagefraction}{0.7}

@@

The 2018 release of the SMT Library contained \num{336844}
non-incremental benchmarks in 51 logics, as well as~\num{10167}
incremental benchmarks (with a total of nearly 32 million {\tt
  check-sat} commands) in 26 logics.  Table~\ref{table:smtlib}
presents further data on the size of the SMT Library since~2015.
While the library usually grows from one year to the other because of
the inclusion of new benchmarks and logics, continuous curation
efforts may also cause benchmarks to be removed or reclassified on
occasion, thereby causing individual logics or even the entire library
to shrink.  From~2015 to~2018, the total number of benchmarks has
increased by~68\,\%.  The largest benchmark in the 2018 release had a
size of \SI{1041}{MB}; the total size of the release was \SI{91}{GB}.

@@

Jobs consist of individual \emph{job pairs}.  Each job pair
corresponds to a specific solver/benchmark combination.
Figure~\ref{fig:job-pairs} shows the total number of job pairs for the
SMT Competition from~2015 to~2018.  Over this period, the number has
grown by~72\,\%, from~\num{1028615} job pairs in~2015 to~\num{1776062}
job pairs in~2018.  (This is the number of unique solver/benchmark
combinations that were part of the competition.  The number of job
pairs that were actually run is even higher: test jobs and job pairs
that had to be re-run, e.g., because of issues with the StarExec
infrastructure, are not included.)  For comparison, the~2014
competition only ran~\num{339714} job pairs~\cite{CDW14}.

@@

Participants were not required to provide source code with their
submission.  There were no licensing requirements/restrictions for the
submitted binaries of the participating solvers.
%
Participants were encouraged to submit a short system description of
their solver, including a list of its authors and an explanation of
the SMT solving approach used.  System descriptions are published on
the competition website.  They were submitted for about 60\,\% of all
solver versions since 2015.

@@

\include{table-participants}

@@

The competition-wide ranking for the \maintrack was retained and
refined after~2014.  Sequential and parallel performance are reported
separately.  Let $N_i$ be the total number of benchmarks in
division~$i$ that were used in the competition (and not removed
because of disagreements), and let $\langle e_i, n_i, w_i, c_i\rangle$
be a solver's raw score for this division.
%
In 2015 and~2016, the \emph{competition-wide raw score} for a solver
was defined as
%
$$\sum_i (e_i == 0 \,?\, (n_i/N_i)^2 : -e_i) \ln N_i$$
%
where the sum is over all competitive divisions into which the solver
was entered.
%
In 2017, the (per-division) penalty for erroneous results was changed
from the number of errors to a fixed constant, since the division
weight~$\ln N_i$ already takes the number of benchmarks in the
division into account, and multiplying it by the number of errors
overemphasizes large divisions.  In 2017 and 2018, the
\emph{competition-wide raw score} for a solver was thus defined as
%
$$\sum_i (e_i == 0 \,?\, (n_i/N_i)^2 : -4) \ln N_i$$
%
The value~$-4$ was chosen to balance the strong interest in correct
solvers against the risk of stifling innovation.  In the division
rankings (Section~\ref{sec:scoring-division-rankings}), an incorrect
response causes a solver to be ranked below all correct solvers for
that division.  But the competition's \maintrack consists of many
independent divisions; a solver bug that affects one division need not
manifest itself in other divisions.  With this value, entering a
(possibly buggy) solver that can solve all benchmarks into a division
has a positive expected score if the probability that a soundness bug
in the solver will be triggered is below~$20\,\%$.

@@

\include{table-winners-main}

\include{table-winners-application}

\include{table-winners-unsat-core}

@@

The 2018 edition of the competition was part of the FLoC Olympic
Games, which sponsored seven medals.  The competition organizers
awarded the following solvers:
\begin{itemize}
\item The best three \maintrack solvers according to the
  competition-wide ranking (CVC4, Yices, and SMTInterpol; see
  Table~\ref{table:competition-wide-ranking}),
\item the solvers that won the QF\_BV division in the \maintrack
  (Boolector and MinkeyRink; see Table~\ref{tab:winners-main}),
\item the solver that won the most competitive divisions in the
  \apptrack (Yices; see Table~\ref{tab:winners-application}),
\item the solver that won the most competitive divisions in the
  \ucoretrack (Yices; see Table~\ref{tab:winners-unsat-core}).
\end{itemize}

@@

\begin{figure}
  \begin{subfigure}[t]{0.5\textwidth}
    \scalebox{.35}{\includegraphics{%
        plot/pdf/cactus_plot-Virtual_Best_Solver_(VBS)_Main_Track_2014-2015.pdf}}
    \caption{2014 vs.~2015}
  \end{subfigure}
  \begin{subfigure}[t]{0.5\textwidth}
    \scalebox{.35}{\includegraphics{%
      plot/pdf/cactus_plot-Virtual_Best_Solver_(VBS)_Main_Track_2015-2016.pdf}}
    \caption{2015 vs.~2016}
  \end{subfigure}
  \\[2ex]
  \begin{subfigure}[t]{0.5\textwidth}
    \scalebox{.35}{\includegraphics{%
      plot/pdf/cactus_plot-Virtual_Best_Solver_(VBS)_Main_Track_2016-2017.pdf}}
    \caption{2016 vs.~2017}
  \end{subfigure}
  \begin{subfigure}[t]{0.5\textwidth}
    \scalebox{.35}{\includegraphics{%
    plot/pdf/cactus_plot-Virtual_Best_Solver_(VBS)_Main_Track_2017-2018.pdf}}
    \caption{2017 vs.~2018}
  \end{subfigure}
  \caption{VBS on common benchmarks for pairs of years of all \maintrack
  divisions with a time limit of \SI{1200}{s}.  Number of common benchmarks is
  indicated with a gray vertical line.}
  \label{fig:progress}
\end{figure}

@@

For the \maintrack, we determine the VBS as the best wall-clock
performance per (correctly solved) instance with the time limit of
\SI{1200}{s} used in the competitions 2017 and 2018.
Figure~\ref{fig:progress-all} visualizes VBS performance on common
benchmarks from 2015--2018 as cactus plot.  Figure~\ref{fig:progress}
shows VBS performance on common benchmarks over pairs of consecutive
years from 2014 to 2018.  Both plots show the wall-clock solving time
per instance over all instances, sorted by solving time.
%
Between 2014 (\num{67426} benchmarks) and~2015 (\num{154238}
benchmarks), the number of common benchmarks is \num{67070}; between
2015 and 2016 (\num{154424} benchmarks) it is \num{154238}; between
2016 and 2017 (\num{238758} benchmarks) it is \num{145236}; and
between 2017 and 2018 (\num{333241} benchmarks) it is \num{226429}.
The number of solved instances on these sets of common benchmarks
increased from 2014 to 2015 by 0.1\%, from 2016 to 2017 by 3.9\%, and
from 2017 to 2018 by 0.3\%.  Interestingly, it decreased from 2015 to
2016 by 3.5\%.  This is mainly due to the fact that, in 2015, more
than \num{5600} benchmarks of division QF\_FP were solved by one of
the two participating configurations of Z3 but by no other solver, and
in~2016 Z3 did not participate in this division.

For the \apptrack, we determine the VBS as the best performance in
terms of number of correct \texttt{check-sat} answers within the time
limit of \SI{2400}{s} (the time limit of the \apptrack in the
competitions 2015--2018).  The total number of eligible
\texttt{check-sat} queries (i.e., all queries with status sat or unsat
up to the first query with status unknown) on common benchmarks from
2015--2018 is \num{30370517}.  Figure~\ref{fig:progress-all-app}
visualizes VBS performance on these benchmarks as a bar plot.  In
2015, already 99.95\% of the eligible queries were answered correctly,
which naturally does not leave a lot of room for improvement.  In
2016, performance decreased by \num{185625} correct answers (0.61\%),
which can be entirely attributed to a performance regression of Z3 on
two benchmarks in the QF\_NIA division.

>>>>

% float layout
\setcounter{topnumber}{2}
\setcounter{bottomnumber}{2}
\setcounter{totalnumber}{4}
\renewcommand{\topfraction}{0.85}
\renewcommand{\bottomfraction}{0.85}
\renewcommand{\textfraction}{0.15}
\renewcommand{\floatpagefraction}{0.75}

@@

The 2018 release of the SMT Library contained \num{336844}
non-incremental benchmarks in 51 logics, as well as~\num{10167}
incremental benchmarks (with a total of nearly 32 million {\tt
  check-sat} commands) in 26 logics.  Table~\ref{table:smtlib}
presents further data on the size of the SMT Library since~2015.
While the library usually grows from one year to the other because of
the inclusion of new benchmarks and logics, continuous curation
efforts may also cause benchmarks to be removed or reclassified on
occasion, thereby causing individual logics or even the entire library
to shrink.  From~2015 to~2018, the total number of benchmarks has
increased by~\SI{68}{\percent}.  The largest benchmark in the 2018
release had a size of \SI{1041}{MB}; the total size of the release was
\SI{91}{GB}.

@@

Jobs consist of individual \emph{job pairs}.  Each job pair
corresponds to a specific solver/benchmark combination.
Figure~\ref{fig:job-pairs} shows the total number of job pairs for the
SMT Competition from~2015 to~2018.  Over this period, the number has
grown by~\SI{72}{\percent}, from~\num{1028615} job pairs in~2015
to~\num{1776062} job pairs in~2018.  (This is the number of unique
solver/benchmark combinations that were part of the competition.  The
number of job pairs that were actually run is even higher: test jobs
and job pairs that had to be re-run, e.g., because of issues with the
StarExec infrastructure, are not included.)  For comparison, the~2014
competition only ran~\num{339714} job pairs~\cite{CDW14}.

@@

Participants were not required to provide source code with their
submission.  There were no licensing requirements/restrictions for the
submitted binaries of the participating solvers.
%
Participants were encouraged to submit a short system description of
their solver, including a list of its authors and an explanation of
the SMT solving approach used.  System descriptions are published on
the competition website.  They were submitted for about
\SI{60}{\percent} of all solver versions since 2015.

@@

\input{table-participants}

@@

The competition-wide ranking for the \maintrack was retained and
refined after~2014.  Sequential and parallel performance are reported
separately.  Let $N_i$ be the total number of benchmarks in
division~$i$ that were used in the competition (and not removed
because of disagreements), and let $\langle e_i, n_i, w_i, c_i\rangle$
be a solver's raw score for this division.
%
In 2015 and~2016, the \emph{competition-wide raw score} for a solver
was defined as
%
$$\sum_i (e_i == 0 \,?\, (n_i/N_i)^2 : -e_i) \ln N_i$$
%
where the sum is over all competitive divisions into which the solver
was entered.
%
In 2017, the (per-division) penalty for erroneous results was changed
from the number of errors to a fixed constant, since the division
weight~$\ln N_i$ already takes the number of benchmarks in the
division into account, and multiplying it by the number of errors
overemphasizes large divisions.  In 2017 and 2018, the
\emph{competition-wide raw score} for a solver was thus defined as
%
$$\sum_i (e_i == 0 \,?\, (n_i/N_i)^2 : -4) \ln N_i$$
%
The value~$-4$ was chosen to balance the strong interest in correct
solvers against the risk of stifling innovation.  In the division
rankings (Section~\ref{sec:scoring-division-rankings}), an incorrect
response causes a solver to be ranked below all correct solvers for
that division.  But the competition's \maintrack consists of many
independent divisions; a solver bug that affects one division need not
manifest itself in other divisions.  With this value, entering a
(possibly buggy) solver that can solve all benchmarks into a division
has a positive expected score if the probability that a soundness bug
in the solver will be triggered is below~\SI{20}{\percent}.

@@

\input{table-winners-main}

\input{table-winners-application}

\input{table-winners-unsat-core}

@@

The 2018 edition of the competition was part of the FLoC Olympic
Games, which sponsored seven medals.  The competition organizers
awarded the following solvers:
\begin{itemize}
\item The best three \maintrack solvers according to the
  competition-wide ranking (CVC4, Yices, and SMTInterpol; see
  Table~\ref{table:competition-wide-ranking}),
\item the solvers that won the QF\_BV division in the \maintrack
  (Boolector and Minkey\-Rink; see Table~\ref{tab:winners-main}),
\item the solver that won the most competitive divisions in the
  \apptrack (Yices; see Table~\ref{tab:winners-application}),
\item the solver that won the most competitive divisions in the
  \ucoretrack (Yices; see Table~\ref{tab:winners-unsat-core}).
\end{itemize}

@@

\begin{figure}
  \begin{subfigure}[t]{0.5\textwidth}
    \scalebox{.35}{\includegraphics{%
        plot/pdf/cactus_plot-Virtual_Best_Solver_(VBS)_Main_Track_2014-2015.pdf}}
    \caption{2014 vs.~2015}
  \end{subfigure}%
  \begin{subfigure}[t]{0.5\textwidth}
    \scalebox{.35}{\includegraphics{%
      plot/pdf/cactus_plot-Virtual_Best_Solver_(VBS)_Main_Track_2015-2016.pdf}}
    \caption{2015 vs.~2016}
  \end{subfigure}
  \\[2ex]
  \begin{subfigure}[t]{0.5\textwidth}
    \scalebox{.35}{\includegraphics{%
      plot/pdf/cactus_plot-Virtual_Best_Solver_(VBS)_Main_Track_2016-2017.pdf}}
    \caption{2016 vs.~2017}
  \end{subfigure}%
  \begin{subfigure}[t]{0.5\textwidth}
    \scalebox{.35}{\includegraphics{%
    plot/pdf/cactus_plot-Virtual_Best_Solver_(VBS)_Main_Track_2017-2018.pdf}}
    \caption{2017 vs.~2018}
  \end{subfigure}
  \caption{VBS on common benchmarks for pairs of years of all \maintrack
  divisions with a time limit of \SI{1200}{s}.  Number of common benchmarks is
  indicated with a gray vertical line.}
  \label{fig:progress}
\end{figure}

@@

For the \maintrack, we determine the VBS as the best wall-clock
performance per (correctly solved) instance with the time limit of
\SI{1200}{s} used in the competitions 2017 and 2018.
Figure~\ref{fig:progress-all} visualizes VBS performance on common
benchmarks from 2015--2018 as cactus plot.  Figure~\ref{fig:progress}
shows VBS performance on common benchmarks over pairs of consecutive
years from 2014 to 2018.  Both plots show the wall-clock solving time
per instance over all instances, sorted by solving time.
%
Between 2014 (\num{67426} benchmarks) and~2015 (\num{154238}
benchmarks), the number of common benchmarks is \num{67070}; between
2015 and 2016 (\num{154424} benchmarks) it is \num{154238}; between
2016 and 2017 (\num{238758} benchmarks) it is \num{145236}; and
between 2017 and 2018 (\num{333241} benchmarks) it is \num{226429}.
The number of solved instances on these sets of common benchmarks
increased from 2014 to 2015 by \SI{0.1}{\percent}, from 2016 to 2017
by \SI{3.9}{\percent}, and from 2017 to 2018 by \SI{0.3}{\percent}.
Interestingly, it decreased from 2015 to 2016 by \SI{3.5}{\percent}.
This is mainly due to the fact that, in 2015, more than \num{5600}
benchmarks of division QF\_FP were solved by one of the two
participating configurations of Z3 but by no other solver, and in~2016
Z3 did not participate in this division.

For the \apptrack, we determine the VBS as the best performance in
terms of number of correct \texttt{check-sat} answers within the time
limit of \SI{2400}{s} (the time limit of the \apptrack in the
competitions 2015--2018).  The total number of eligible
\texttt{check-sat} queries (i.e., all queries with status sat or unsat
up to the first query with status unknown) on common benchmarks from
2015--2018 is \num{30370517}.  Figure~\ref{fig:progress-all-app}
visualizes VBS performance on these benchmarks as a bar plot.  In
2015, already \SI{99.95}{\percent} of the eligible queries were
answered correctly, which naturally does not leave a lot of room for
improvement.  In 2016, performance decreased by \num{185625} correct
answers (\SI{0.61}{\percent}), which can be entirely attributed to a
performance regression of Z3 on two benchmarks in the QF\_NIA
division.
